---
title: "Contrasting between- and within-rows stem profiles in poplar plantations - Modelling tree and 'Treat' effect - HLM approach"
author: "Puletti N., Scotti R., ???"
date: june 2018
output:
  pdf_document:
    toc: true
---

# Access to the data
```{r}
library(tidyverse)
library(magrittr)
library(broom)

source("DataWrangling.R")

```


*** following: https://www.jstatsoft.org/article/view/v056i05/ ***

# Level 1 - 'TreeId'
We fit an initial two-level HLM using crossection height (Sect_height) to explain differences between profiles (delta_d) allowing for a random intercept for each tree:     
***   delta_d ~ Sect_height + (1 | TreeId)   ***

```{r}
library(lme4)

mc <- function(frml, ...) {
# evaluate the model (specified as formula) and produce the residulas control plot
  fm <- stff %>%  lmer(data = ., frml, ... )
  p <- qplot(x = stff$Sect_height, y = residuals(fm), 
        geom = c("point", "smooth")) + ylab("level-1 residuals") +
    ggtitle(paste0(as.expression(frml)))
  print(p)
return(fm)
}

# prova senza Treat e senza Sect_height
fm00 <- mc(delta_d_cm ~  (1 | TreeId))

# prova con Treat ma senza Sect_height
fm10 <- mc(delta_d_cm ~ Treat + (1 | TreeId))
summary(fm10)
# Treat  non ha effetto! (più evidente dopo, con anova)
summary(fm00)

```
## The  two orthogonal profiles are significantly different: 1.8 cm +/- 10%
```{r}
anova(fm10, fm00)

# Contributo di Treat non significativo
qplot(x = stff$Treat, y = residuals(fm10), 
      geom = c("boxplot")) + ylab("level-1 residuals")
# residui distribuiti male e non a zero di media per Treat

```


```{r}
# prova CON Sect_height,   lineare
fm01 <- mc(delta_d_cm ~ Sect_height + (1 | TreeId))
fm11 <- mc(delta_d_cm ~ Treat + Sect_height + (1 | TreeId))
anova(fm11, fm01)
qplot(x = stff$Treat, y = residuals(fm11), 
      geom = c("boxplot")) + ylab("level-1 residuals")
# residui ancora etreoscedastici

# prova CON Sect_height, parabolica
fm12 <- mc(delta_d_cm ~ Treat + (1 | TreeId) + Sect_height *(1 | TreeId) + I(Sect_height^2)* (1 | TreeId))
fm02 <- mc(delta_d_cm ~ (1 | TreeId) + Sect_height *(1 | TreeId) + I(Sect_height^2)* (1 | TreeId))
anova(fm02,fm12)
stff %>% ggplot() +
  geom_point(aes(x = Sect_height, y = delta_d_cm)) +
  geom_smooth(aes(x = Sect_height, y = delta_d_cm)) +
  geom_line(aes(x = Sect_height, y = fitted(fm12)), color = "yellow", size = 1) + 
  facet_wrap(~TreeId)

# prova CON Sect_height, al cubo (con RML=T, non converge!)
fm13 <- mc(delta_d_cm ~ Treat + (1 | TreeId) 
           + Sect_height *(1 | TreeId) 
           + I(Sect_height^2)* (1 | TreeId)
           + I(Sect_height^3)* (1 | TreeId)
           , REML = F)
fm03 <- mc(delta_d_cm ~ (1 | TreeId) 
           + Sect_height *(1 | TreeId) 
           + I(Sect_height^2)* (1 | TreeId)
           + I(Sect_height^3)* (1 | TreeId)
           , REML = F)
anova(fm03,fm13)

# l'effetto di Treat, non si manifesta in nessun modo!!
#  Pr(>Chisq) = 0.761  !! Conferma, non si vede differenza


stff %>% ggplot() +
  geom_point(aes(x = Sect_height, y = delta_d_cm)) +
  geom_smooth(aes(x = Sect_height, y = delta_d_cm)) +
  geom_line(aes(x = Sect_height, y = fitted(fm13)), color = "yellow", size = 1) + 
  facet_wrap(~TreeId)


```

In an exploratory setting we would go through this cycle of residual analysis, identification of explanatory variables, and model fitting multiple times until a satisfactory level-1 model is found (Tukey 1977).

# Level-2 residuals
The level-2, or random effects, residuals are defined as Zibi or, more commonly, bi.
Following an upward residual analysis we prefer the use of EB residuals at level 2 (insted of LS residuals).
The level-2 residuals are used to
* identify additional explanatory variables that contribute significantly to the model,
* check linearity of the level-2 explanatory variables, and
* investigate whether the level-2 residuals follow a normal distribution.
To obtain the level-2 EB residuals from model fm3, we use the following code:
```{r}
resid2_fm13 <- HLMresid(object = fm13, level = "TreeId")
head(resid2_fm13)

res <- stff %>% 
  select(Treat, TreeId) %>% 
  unique() %>% 
  inner_join(mutate(resid2_fm13, TreeId = row.names(resid2_fm12))) %>%
  select(1:2, pw0coeff = 3, pw1coeff = 4, pw2coeff = 5, pw3coeff = 6) %>%
  gather(param, value,  3:6)

# paste0("pw",0:3,"coeff = ", 3:6, collapse=", ")

ggplot(res, aes(x = Treat, y = value)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  facet_wrap(~param, scales = "free")

res %$%
  qplot(x = Treat, y = value, geom = 'boxplot', facets = ~param,
  xlab = 'distance between rows', main = 'level-2 residual')



```



# Opzione non esplorata:
Forse è più semplice dividere il range di Sect_height in 3 parti < 1 e >4 (zeri dell'intepolazione generale dei residui dal modello lineare) ed utilizzare quats avriabile come ulteriore 'effetto fisso'. A quel punto l'approssimazione lineare di delta_d=f(Sect_heiht) dovrebbe essere presentabile.

[rmarkdown::render("PdP_HLM3.Rmd", encoding="UTF-8")]
